---
title: "METRIC advanced"
author: "Guillermo Federico Olmedo"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{METRIC advanced}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

abstract

## Introduction

A brief introduction to water and METRIC

First we have to load water package: 

```{r}
library(water)

```

## Base data preparation

To calculate METRIC Evapotranspiration using water in the adavanced way, you need 
four sources:

- A raw Landsat image
- A Weather Station 
- A polygon with our Area-of-interset SpatialPolygon object, in case we don't 
want to calc ET for all the landsat scene.
- A Digital Elevation model

First we create a polygon from bottomright and topleft point: 
```{r}
aoi <- createAoi(topleft = c(273110, -3914450), 
                 bottomright = c( 288050, -3926650), EPSG = 32619)
```

Then, we load the weather station data. We are going to use function `read.WSdata`. This function converts our csv file into a `waterWeatherStation` object. And, if we provide a Landsat metadata file, will calculate the weather conditions at the 
satellite flyby.

```{r}
csvfile <- system.file("extdata", "apples.csv", package="water")
MTLfile <- system.file("extdata", "L7.MTL.txt", package="water")
WeatherStation <- read.WSdata(WSdata = csvfile, date.format = "%d/%m/%Y",
                  lat=-35.42222, long= -71.38639, elev=201, height= 2.2,
                  MTL = MTLfile)
```

We can visualize our weather station data: 
```{r, fig.width = 5}
print(WeatherStation)

plot(WeatherStation, alldata=FALSE)
```


And after that, we load the image. Usually, when using `water` we can use the 
function `loadImage` to load a Landsat image from `TIF files` download direclty
from [Earth Explorer](http://earthexplorer.usgs.gov/). In this vignette we are
going to use some Landsat 7 example data wich comes with **water** package.
```{r, fig.width = 5}
image.DN <- L7_Talca[[c(1:5,7)]]
B6 <- L7_Talca[[6]]
```

Finally we are going to create the Digital Elevation Model from the needed grid
files from [Earth Explorer](http://earthexplorer.usgs.gov/). First we check wich
grid files we are going to need:

```{r}
checkSRTMgrids(image.DN)
```

You should download all the needed grid files, and then, you can use `prepareSRTMdata(extent = image.DN)` to create the DEM. In this vignette we are going to load the example data provided with **water** package.

```{r}
DEM <- DEM_Talca
```

## Net Radiation

```{r}
surface.model <-METRICtopo(DEM)

solar.angles.r <- solarAngles(surface.model = surface.model, 
                              WeatherStation = WeatherStation, MTL = MTLfile)

Rs.inc <- incSWradiation(surface.model = surface.model, 
                         solar.angles = solar.angles.r, 
                         WeatherStation = WeatherStation)

image.TOAr <- calcTOAr(image.DN = image.DN, sat="L7", MTL = MTLfile, 
                       incidence.rel = solar.angles.r$incidence.rel)

image.SR <- calcSR(image.TOAr=image.TOAr, sat = "L7", 
                   surface.model=surface.model, 
                   incidence.hor = solar.angles.r$incidence.hor, 
                   WeatherStation=WeatherStation, ESPA = F)

albedo <- albedo(image.SR = image.SR,  coeff="Tasumi", sat = "L7")

LAI <- LAI(method = "metric2010", image = image.TOAr, L=0.1, sat = "L7")

Ts <- surfaceTemperature(LAI=LAI, sat = "L7", thermalband = B6,
                         WeatherStation = WeatherStation)

Rl.out <- outLWradiation(LAI = LAI, Ts=Ts)

Rl.inc <- incLWradiation(WeatherStation,DEM = surface.model$DEM, 
                         solar.angles = solar.angles.r, Ts= Ts)

Rn <- netRadiation(LAI, albedo, Rs.inc, Rl.inc, Rl.out)
```

## Soil Heat Flux

```{r}
G <- soilHeatFlux(image = image.SR, Ts=Ts,albedo=albedo, 
                  Rn=Rn, image.SR, LAI=LAI, sat = "L7")
```

## Sensible Heat Flux

```{r, fig.width = 5}
Z.om <- momentumRoughnessLength(LAI=LAI, mountainous = TRUE, 
                                method = "short.crops", 
                                surface.model = surface.model)

hot.and.cold <- calcAnchors(image = image.TOAr, Ts = Ts, LAI = LAI, plots = F,
                            albedo = albedo, Z.om = Z.om, n = 1, 
                            anchors.method = "CITRA-MCB", sat = "L7", 
                            deltaTemp = 5, verbose = FALSE)

H <- calcH(anchors = hot.and.cold, Ts = Ts, Z.om = Z.om, 
           WeatherStation = WeatherStation, ETp.coef = 1.05, sat = "L7", 
           Z.om.ws = 0.0018, DEM = DEM, Rn = Rn, G = G, verbose = FALSE)
```

# Daily Evapotranspiration estimation

```{r, fig.width = 5}
ET.24 <- ET24h(Rn, G, H$H, Ts, WeatherStation, ETr.daily=6.57)
```

